Introduction

### 2015-07-14 dissertation/introdefhistory.tex
limit per resource

At this point, I left Opera to work as a consultant on Semantic Web
technologies. One of the projects that were successfully completed was
called Sublima\cite{TODO}. It was driven entirely by RDF and SPARQL,
and featured faceted navigation, navigation in a thesaurus, and full
text search. Experiences from this project influenced requirements of
SPARQL 1.1, and I was an editor of the SPARQL 1.1 Features and
Requirements specification\cite{TODO}. Parts of the property paths
feature and aggregate queries were used extensively. Write operations
were also implemented with SPARQL using what was then specified only
in a member submission to the W3C\cite{TODO}. This was the basis for
the update language in SPARQL 1.1 but differed substantially in
surface syntax, partly because our work showed how certain patterns
could be simplified. Another project, based on much of the same code
base replaced the SKOS ontology of the Sublima project with an OWL
ontology, and also used reasoning to aid navigation in multimedia
libraries.

>>>>

At this point, I left Opera to work as a consultant on Semantic Web
technologies. One of the projects that were successfully completed was
called Sublima\cite{TODO}. It was driven entirely by RDF and SPARQL,
and featured faceted navigation, navigation in a thesaurus, and full
text search. Experiences from this project influenced requirements of
SPARQL 1.1, and I was an editor of the SPARQL 1.1 Features and
Requirements specification\cite{TODO}. Parts of the property paths
feature and aggregate queries were used extensively. Write operations
were also implemented with SPARQL using what was then specified only
in a member submission to the W3C\cite{TODO}. This was the basis for
the update language in SPARQL 1.1 but differed substantially in
surface syntax, partly because our work showed how certain patterns
could be simplified.

Two features that arose from Sublima requirements were not accepted
for SPARQL 1.1, however: Full text index and a feature known as
``Limit Per Resource''\footnote{see
  \href{http://www.w3.org/2009/sparql/wiki/Feature:LimitPerResource}}. The
latter feature arises from the fact that usually, the user is
indifferent to the number of solutions to a query, and therefore
limiting by the number of solutions is unhelpful. This is especially
true since other features of SPARQL, such as optional clauses, are
helpful in dealing with heterogeneous data. In the Sublima case, it
was interesting to limit by the number of articles that were returned,
but the number of solutions depended on the number of authors of an
article and the concepts used to classify an article. To achieve the
desired effect with SPARQL 1.1, one would have to write a very complex
subquery, which is undesirable for such an important feature.

Another project, based on much of the same code
base replaced the SKOS ontology of the Sublima project with an OWL
ontology, and also used reasoning to aid navigation in multimedia
libraries.
